---
title: "**Agreement of MARCS `r gsub(x=gsub(x=params$this_endpoint, pattern='All_', replacement='All (tertile) '), pattern='_', replacement=' ')` with Path Phenotypes in `r gsub(x=params$study, pattern='_', replacement=' ')`**"
author: "MARCS Crossfunctional Team"
date: last-modified
date-format: "MMM D, YYYY   hh:mm"
format: !include ../shared-templates/standard-header.yml
params: !include ../shared-templates/standard-params.yml
execute: !include ../shared-templates/standard-execute.yml
editor_options: 
  chunk_output_type: console
---

\clearpage

The Takeda MARCS VCE tool is a semi-automated, computer-assisted small intestine villous damage scoring device that provides an effective way to implement the per-frame CE-VAST scale and the per-region MARCS, including the calculation of the individual biomarker outputs, notably BVA using white-light video VCE examinations of patients. The use of this tool is contingent upon the availability of images obtained by a compatible capsule endoscopy device. The tool includes a marking capability that inputs VCE videos and defines regions, which output may then be used by either a manual scoring tool that facilitates manual scoring, or a machine-learning algorithm (with static code) to automate scoring. 

```{=tex}
\begin{figure}
  \caption{Schematic Depiction of the SI and Areas Evaluated by MARCS}
  \centering
  \includegraphics{images/BVA.png}
  \caption*{Burden of Villous Atrophy (BVA) is the mean of the MARCS curve in Tertile 1.}
\end{figure}
```

Validation will be conducted in multiple clinical trials. Specifically, beginning with (i) evaluation of the CE-VAST scale on a per-frame basis, (ii) analytical validation of BVA, and (iii) preliminary assessment of clinical/biological validation, including the analysis on the following study data:

```{=tex}
\begin{figure}
  \caption{Study Calendar}
  \centering
  \includegraphics{images/Study_Design.png}
  \caption*{Scheduled visits identify when data of various types is collected.}
\end{figure}
```

The analysis documented here is also performed on other studies undertaken by the franchise.

\clearpage

```{r analysis, echo=FALSE, warning=FALSE, error=FALSE}
source(file.path(params$repo_clone, "Shared-Content", "setup-analysis-environment.R"))
source(file.path(params$repo_clone, "Shared-Content", "purrr-helpers.R"))
analysis_ctx <- setup_analysis_environment(params, "CE-VAST-Agreement")

form_merged_data_sets_output <- form_merged_data_sets(analysis_type=params$analysis_type, study=study, this_endpoint=this_endpoint, output_dir=output_dir)

# Custom data preparation for CE-VAST Agreement analysis
prepare_cevast_data <- function(merged_data_set, label_name) {
  # Only use ML data for this analysis
  merged_data_set$merged_data.ML$merged_data.ML.full
}

# Custom analysis wrapper to handle CE-VAST specific requirements
cevast_analysis_wrapper <- function(analysis_data, reader_type = "ML") {
  results <- main_analysis(analysis.data.in = analysis_data, reader_type = reader_type)
  
  # Save extended analysis data to CSV as required
  write.csv(results$analysis.data.in_ext, 
           path(output_dir, paste0("8 analysis.data.in_ext-", reader_type, "-", 
                                   analysis_data$label.name[1], ".csv")), 
           row.names = FALSE)
  
  return(results)
}

# Execute analysis using functional approach for ML data only
ml_results <- process_merged_data_sets(
  form_merged_data_sets_output,
  cevast_analysis_wrapper,
  prepare_cevast_data,
  reader_type = "ML"
)

# Extract results into the expected named lists for downstream processing
kappa.report.ML <- map(ml_results, ~ .x$kappa.report)
analysis.data.in_ext.ML <- map(ml_results, ~ .x$analysis.data.in_ext)
kappaplot.ML <- map(ml_results, ~ .x$kappaplot)

# Extract stratified results (only if they exist)
kappa.report.ML.byRandStat <- map(ml_results, ~ .x$kappa.report.byRandStat) %>% 
  discard(is.null)
kappaplot.ML.byRandStat <- map(ml_results, ~ .x$kappaplot.byRandStat) %>% 
  discard(is.null)

kappa.report.ML.byPairing <- map(ml_results, ~ .x$kappa.report.byPairing) %>% 
  discard(is.null)
kappaplot.ML.byPairing <- map(ml_results, ~ .x$kappaplot.byPairing) %>% 
  discard(is.null)

kappa.report.ML.byArm <- map(ml_results, ~ .x$kappa.report.byArm) %>% 
  discard(is.null)
kappaplot.ML.byArm <- map(ml_results, ~ .x$kappaplot.byArm) %>% 
  discard(is.null)

# Initialize empty Human lists (not used in this analysis but expected by downstream code)
kappa.report.Human <- list()
kappa.report.Human.byRandStat <- list()
kappa.report.Human.byPairing <- list()
kappa.report.Human.byArm <- list()
kappaplot.Human <- list()
analysis.data.in_ext.Human <- list()
kappaplot.Human.byRandStat <- list()
kappaplot.Human.byPairing <- list()
kappaplot.Human.byArm <- list()
```

## Executive Summary for Machine Learning Reads

We assess the agreement between histology-determined phenotypes by Esophagogastroduodenoscopy (EGD) biopsy samples and MARCS score derived from VCE videos. This agreement is estimated using Cohen's kappa coefficient ($\kappa$, lowercase Greek kappa), a statistic that is used for qualitative (categorical) items in its weighted form to take into account greater penalty for larger disagreement (that is, calling a class which is very different from the reference class as opposed to calling a class which is nearer).

The MARCS score used here is determined from the first 5% of the segmented video, representing the region routinely biopsied with EGD. The MARCS score is on the CE-VAST scale, resulting in the weighted $\kappa$ to evaluate the ability of CE-VAST to align with pathologist-determined Marsh-Oberhuber type. Measures of agreement are also presented for HLA_DQ, HLA_G phenotypes if available in the study data.

```{r pathologist-agreement-executive-summary.ML, echo=FALSE}
generate_kappa_exec_summary(kappa.summary.in=kappa.report.ML, reader_type="ML")
```

```{r kappa_exec_fig.ML, echo=FALSE}
forPDF <- NULL
forPDF <- generate_kappa_fig(kappa.report.ML, reader_type="ML")
if (!is.null(forPDF)) {
  ggsave(forPDF, filename=path(figure_path, "pathologist-agreement-executive-summary.ML.png"))
  forPDF
}
```

\clearpage
## Stratified Analyses

### by Randomization
```{r agreement-byRandomization.ML, echo=FALSE}
byRandomization <- NULL
byRandomization <- generate_kappa_bystratum(kappa.summary.in=kappa.report.ML.byRandStat, reader_type="ML", stratifier="RandStat")
if (!is.null(byRandomization)) {
  byRandomization
}
```
```{r patholoagreement-byRandomization.ML.fig, echo=FALSE}
forPDF <- NULL
forPDF <- generate_kappa_fig_bystratum(kappa.summary.in=kappa.report.ML.byRandStat, reader_type="ML", stratifier="RandStat")
if (!is.null(forPDF)) {
  forPDF
}
```
```{r agreement-byRandomization-empty, echo=FALSE}
if (is.null(byRandomization)) {
  cat("Stratification by randomization status not possible on the basis of available study data.")
}
```

\clearpage
### by Visit
```{r pathologist-agreement-byPairing.ML, echo=FALSE}
byPairing <- NULL
byPairing <- generate_kappa_bystratum(kappa.summary.in=kappa.report.ML.byPairing, reader_type="ML", stratifier="Pairing")
if (!is.null(byPairing)) {
  byPairing
}
```
```{r pathologist-agreement-byPairing.ML.fig, echo=FALSE}
forPDF <- NULL
forPDF <- generate_kappa_fig_bystratum(kappa.summary.in=kappa.report.ML.byPairing, reader_type="ML", stratifier="Pairing")
if (!is.null(forPDF)) {
  forPDF
}
```
```{r pathologist-agreement-byPairing.ML-empty, echo=FALSE}
if (is.null(byPairing)) {
  cat("Stratification by visit not possible on the basis of available study data.")
}
```

\clearpage
### by Arm
```{r agreement-byArm.ML, echo=FALSE}
byArm <- NULL
byArm <- generate_kappa_bystratum(kappa.summary.in=kappa.report.ML.byArm, reader_type="ML", stratifier="Arm")
if (!is.null(byArm)) {
  byArm
}
```
```{r pathologist-agreement-byArm.ML.fig, echo=FALSE}
forPDF <- NULL
forPDF <- generate_kappa_fig_bystratum(kappa.summary.in=kappa.report.ML.byArm, reader_type="ML", stratifier="Arm")
if (!is.null(forPDF)) {
  forPDF
}
```
```{r pathologist-agreement-byArm.ML-empty, echo=FALSE}
if (is.null(byArm)) {
  cat("Stratification by arm not possible on the basis of available study data.")
}
```

<!-- 
NO HUMAN FOR NOW \clearpage
## Details for Human Reads
```{r pathologist-agreement-executive-summary.Human, echo=FALSE}
exec_summary.Human <- NULL
exec_summary.Human <- generate_kappa_exec_summary(kappa.summary.in=kappa.report.Human, reader_type="Human")
if (!is.null(exec_summary.Human)) {
  exec_summary.Human
}
```
```{r kappa_exec_fig.Human, echo=FALSE}
fig.out <- generate_kappa_fig(kappa.summary.in=kappa.report.Human, reader_type="Human")
if (!is.null(fig.out)) {
  fig.out
}
```{r pathologist-agreement-executive-summary-Human-empty, echo=FALSE}
if (is.null(exec_summary.Human)) {
  cat("Human reader analyses not possible on the basis of available study data.")
}
``` 
-->

\clearpage
## Demographic Make-up for Subjects in this Slice

```{r subject-counts, echo=FALSE}
write.csv(form_merged_data_sets_output$counts, file=path(table_path, "subject_counts.csv"))
subject.counts <- form_merged_data_sets_output$counts %>%
  gt() %>%
  tab_header(title="Counts of Subjects and Observations")
gt::gtsave(subject.counts, filename=path(table_path, "subject_counts.formatted.png"))
subject.counts
```
```{r demographics, echo=FALSE}
form_merged_data_sets_output$demographics %>% as_gt() %>%
  cols_width(everything() ~ px(90))
```

### Views to Support Actionable Insights
```{r relationship.plot, fig.width=6, fig.height=8, echo=FALSE}
generate_kappa_relationship_plot(analysis.data.in_ext.ML[[1]], "ML", this_endpoint)
```

\clearpage

## Appendix: Provenance {.unnumbered}
```{r provenance, fig.align='center', echo=FALSE}
cat(paste("Analysis executed by", system2("whoami", stdout=TRUE, stderr=TRUE)))
cat(str_wrap(paste("Results derived from", form_merged_data_sets_output$provenance), width=50, whitespace_only=FALSE))
setwd(repo_clone)
cat(str_wrap(paste("Code used for analysis was", git2r::status(repo=".", staged=TRUE, unstaged=TRUE, untracked=TRUE, ignored=FALSE, all_untracked=FALSE)), width=50, whitespace_only=FALSE))
```
