---
title: "MIRT practice"
format: 
  pdf:
     fig-numbering: FALSE
     include-in-header:  
        - text: |
            \usepackage{pdflscape}
toc: TRUE
number-sections: TRUE
execute: 
  echo: FALSE
  message: FALSE
  warning: FALSE
  tbl-cap-location: top
editor: visual
params:
  printcode: FALSE
---

TEST CHANGE

## Source functions

```{r}
require(psych)
require(tidyverse)
# Note last line here throws an error - but this does not occur when 
source("C:/Users/eri7441/OneDrive - Takeda/Documents/R-Projects/MARCS/Analytical-Validation/Accuracy/functions/accuracy-setup.R")
source("C:/Users/eri7441/OneDrive - Takeda/Documents/R-Projects/MARCS/Shared-Content/preamble.R")
source("C:/Users/eri7441/OneDrive - Takeda/Documents/R-Projects/MARCS/Shared-Content/IRT-models.R")
```

## Build data sets

```{r}
source("C:/Users/eri7441/OneDrive - Takeda/Documents/R-Projects/MARCS/Shared-Content/prep-data-062-2001-histology.R")
source("C:/Users/eri7441/OneDrive - Takeda/Documents/R-Projects/MARCS/Shared-Content/prep-data-062-2001-GISS.R")
```

```{r}
# Visit is ignored
  demog_df <- demog_062_2001_etl()
  histology_df <- histology_062_2001_etl()
  symptoms_df <- symptoms_062_2001_etl()  
```

-   PCA and Factor analysis do not require normality assumption.
-   both suggest centering and scaling first

```{r}
# Adding recipricol. Centering and scaling. Removing rows with missing values of GISS
my.df.working <- my.df %>% mutate(CDVH = 1/VHCD,
                          VHCD.scaled = (VHCD - mean(VHCD, na.rm=T))/sd(VHCD, na.rm=T),
                          CDVH.scaled = (CDVH - mean(CDVH, na.rm=T))/sd(CDVH, na.rm=T),
                          IEL.scaled = (IEL- mean(IEL))/sd(IEL, na.rm=T),
                          GISS.scaled = (GISS - mean(GISS, na.rm=T))/sd(GISS, na.rm=T)) %>% 
  dplyr::filter(is.na(GISS) == FALSE)

my.df.scaled.VHCD <- my.df.working %>% dplyr::select(SUBJID, contains("scaled"), -contains("CDVH"))
names(my.df.scaled.VHCD) <- gsub(x = names(my.df.scaled.VHCD), pattern = ".scaled", replacement = "")

my.df.scaled.CDVH <- my.df.working %>% dplyr::select(SUBJID, contains("scaled"), -contains("VHCD"))
names(my.df.scaled.CDVH) <- gsub(x=names(my.df.scaled.CDVH) , pattern=".scaled", replacement="")
```

## CD:VH deviates from normality

```{r}
hist(my.df.working$VHCD)
hist(my.df.working$CDVH)
hist(log(my.df.working$CDVH))  # Still exhibits skewness

qqnorm(y=my.df.working$VHCD);
qqnorm(y=my.df.working$CDVH);
```

## Factor Analysis

Factor analysis conducted with (VHCD, IEL, GISS) and (CDVH, IEL, GISS). In both cases, negative degrees of freedom and root mean square of the residuals (RMSR) of 0 suggest overfitting. Moreover, FA is typically employed when many variables are associated with a factor. This said, Factor loadings show histology group one one factor and GISS on the another.

### with vh:cd

```{r}
fa_result.vhcd <- fa(my.df.scaled.VHCD %>% dplyr::select(-SUBJID), nfactors = 2, rotate = "varimax")
factor_scores.vhcd <- data.frame(fa_result.vhcd$scores)
names(factor_scores.vhcd) <- c("MR1.VHCD", "MR2.VHCD")
loadings.vhcd <- fa_result.vhcd$loadings
fa_result.vhcd

my.df.working <- my.df.working %>% cbind(factor_scores.vhcd) %>% mutate(DB.FA.VHCD = MR1.VHCD + MR2.VHCD)
```

Health Scores: Higher Scores are associated with high better VHCD, lower IEL Counts, and Lower GISS. Very high correlation with Disease burden metric and Vh:Cd.

```{r}
ggpairs(my.df.working %>% dplyr::select(c(2:4, 10:12)))
```

```{r}
fa_result.cdvh <- fa(my.df.scaled.CDVH %>% dplyr::select(-SUBJID), nfactors = 2, rotate = "varimax")
factor_scores.cdvh <- data.frame(fa_result.cdvh$scores)
names(factor_scores.cdvh) <- c("MR1.CDVH", "MR2.CDVH")
loadings.cdvh <- fa_result.cdvh$loadings
fa_result.cdvh

my.df.working <- my.df.working %>% cbind(factor_scores.cdvh) %>% mutate(DB.FA.CDVH = MR1.CDVH + MR2.CDVH)

ggpairs(my.df.working %>% dplyr::select(c(2:4, 13:15)))
```

Disease burden metrics based on CD:VH have right skewness

```{r}
plot(my.df.working$DB.FA.VHCD, my.df.working$DB.FA.CDVH)
hist(my.df.working$DB.FA.VHCD, breaks=50)
hist(my.df.working$DB.FA.CDVH, breaks=50)
```

## PCA

### PCA with VHCD

Loadings show Histology grouping on PC1 and GISS on PC2

```{r}
# Kaiser-Meyer-Olkin factor adequacy recommended to be larger than 0.6.  We have 0.5
KMO(cor(my.df.scaled.VHCD %>% dplyr::select(-SUBJID)))
# This works out
cortest.bartlett(cor(my.df.scaled.VHCD %>% dplyr::select(-SUBJID)), n = nrow(my.df))

pca_result <- prcomp(x=as.matrix(my.df.scaled.VHCD %>% dplyr::select(-SUBJID)))
pca_result$x <- data.frame(pca_result$x )
names(pca_result$x) <- paste0(names(pca_result$x), ".VHCD")
summary(pca_result)
# eigenvalues
(pca_result$sdev)^2
print(pca_result$rotation)
# Extract loadings
loadings <- pca_result$rotation

my.df.working <- my.df.working %>% cbind(pca_result$x[, 1:2]) %>% mutate(DB.PCA.VHCD = (PC1.VHCD+PC2.VHCD))

```

```{r}
ggpairs(my.df.working %>% dplyr::select(c(2:4, 16:18)))
```

### PCA with CDVH

```{r}
# Kaiser-Meyer-Olkin factor adequacy recommended to be larger than 0.6.  We have 0.5
KMO(cor(my.df.scaled.CDVH %>% dplyr::select(-SUBJID)))
# This works out
cortest.bartlett(cor(my.df.scaled.CDVH %>% dplyr::select(-SUBJID)), n = nrow(my.df))

pca_result <- prcomp(x=as.matrix(my.df.scaled.CDVH %>% dplyr::select(-SUBJID)))
pca_result$x <- data.frame(pca_result$x )
names(pca_result$x) <- paste0(names(pca_result$x), ".CDVH")
summary(pca_result)
# eigenvalues
(pca_result$sdev)^2
print(pca_result$rotation)
# Extract loadings
loadings <- pca_result$rotation

my.df.working <- my.df.working %>% cbind(pca_result$x[, 1:2]) %>% mutate(DB.PCA.CDVH = (PC1.CDVH+PC2.CDVH))

ggpairs(my.df.working %>% dplyr::select(c(2:4, 19:21)))
```

The Disease burden scores using PCA with CD:VH are left skewed and asymmetric.

## IRT

Unidimensionality is not met. This suggests GISS is measuring something different than histology. Here we attempt an IRT Analysis using CDVH. Resulting Patient burden scores exhibit a small range and heaping.

```{r}
IRT.CDVH <- common.alpha.model(data.in = my.df.scaled.CDVH)

PostEst.out.IRT.CDVH <- PosteriorEstimates(data.in = my.df.scaled.CDVH,
  grid = seq(-3, 5, .1),
  I = 3,
  indicator = "C",
  lin = "lin",
  parmatrest = IRT.CDVH$parmatrix,
  stdest = IRT.CDVH$stdmixt)

my.df.working <- my.df.working %>% cbind(PostEst.out.IRT.CDVH %>% dplyr::select(esttheta))

ggpairs(my.df.working %>% dplyr::select(5, 3, 4, esttheta))
hist(my.df.working$esttheta, breaks=50)
```

```{r}
my.df.working %>% ggplot(aes(x=DB.FA.VHCD, y= DB.FA.CDVH)) + geom_point()
my.df.working %>% ggplot(aes(x=DB.FA.VHCD, y= DB.PCA.VHCD)) + geom_point()
my.df.working %>% ggplot(aes(x=DB.PCA.VHCD, y= DB.PCA.CDVH)) + geom_point()
my.df.working %>% ggplot(aes(DB.PCA.VHCD, esttheta))+ geom_point()
```

## MIRT

The mirt package only deals with analysis of dichotomous and polytomous response data using unidimensional and multidimensional latent trait models under the Item Response Theory (IRT) paradigm. This is not a viable option.
