---
title: "Untitled"
format: html
editor: visual
---

## Effect Size

$$d=\frac{\bar{x}_{1} -\bar{x}_{2}}{s} $$

$$ s = \sqrt{\frac{(n_1-1)s_1^2 + (n_2-1)s_2^2}{n_1+n_2-2}}$$

1.  Sample $\bar{x}_{1}$ and $\bar{x}_{2}$ from prior
2.  Sample s from prior
3.  Compute d
4.  Repeat many times to determine proportion of times
    -   d \>= .7
    -   d in .5-.69
    -   d in .35-.49
    -   d \< .35

```{r}
require(mvtnorm)
require(tidyverse)
```

```{r}
my.mean <- c(.5, .15)
my.diag <- rbind(c(.15^2, 0), c(0, .025^2))
my.corr <- rbind(c(1, .25),c(.25, 1))
my.sigma <- sqrt(my.diag) %*% my.corr %*% sqrt(my.diag)


my.sample <- data.frame(rmvnorm(10000, my.mean, my.sigma))
names(my.sample) <- c("Biomarker", "Symptom")

my.sample <- my.sample %>% mutate(Biomarker.result = case_when(Biomarker < .35 ~ 0,
                                                  Biomarker >=.35 & Biomarker < .50 ~ 1,
                                                  Biomarker >=.50 & Biomarker < .70 ~ 2,
                                                  TRUE >= .70 ~ 3),
                     Symptom.result = case_when(Symptom < 0.1 ~ 0,
                                                Symptom >= 0.1 & Symptom < 0.2 ~ 1,
                                                TRUE ~ 2)) %>% 
  mutate(result = case_when(
    Biomarker.result == 3 & Symptom.result != 0 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 2 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 1 ~ "Yellow",
    Biomarker.result == 1 & Symptom.result != 0 ~ "Yellow", 
    .default = "Orange"
  )) 

my.sample %>% ggplot(aes(x=Biomarker, y= Symptom, color=result)) + geom_point() + 
  geom_vline(xintercept = c(.35, .5, .7))+
  geom_hline(yintercept = c(.1, .2))+
  scale_color_manual(values = c("green", "orange", "yellow")) 


anno <- my.sample %>% mutate(Biomarker.result = case_when(Biomarker < .35 ~ 0,
                                                  Biomarker >=.35 & Biomarker < .50 ~ 1,
                                                  Biomarker >=.50 & Biomarker < .70 ~ 2,
                                                  TRUE >= .70 ~ 3),
                     Symptom.result = case_when(Symptom < 0.1 ~ 0,
                                                Symptom >= 0.1 & Symptom < 0.2 ~ 1,
                                                TRUE ~ 2)) %>% 
  mutate(result = case_when(
    Biomarker.result == 3 & Symptom.result != 0 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 2 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 1 ~ "Yellow",
    Biomarker.result == 1 & Symptom.result != 0 ~ "Yellow", 
    .default = "Orange"
  )) %>% count(result) %>% mutate(prop = n/sum(n)) %>% select(-n) 
  
anno$result <- factor(anno$result, "Green", "Yellow", "Orange")

my.sample %>% ggplot(aes(x=Biomarker, y= Symptom, color=result)) + geom_point() + 
  geom_vline(xintercept = c(.35, .5, .7))+
  geom_hline(yintercept = c(.1, .2))+
  scale_color_manual(values = c("green", "orange", "yellow")) + 
  labs(caption = paste("Green: ", anno$prop[1], "\nYellow: ", anno$prop[2], "\nOrange: ", anno$prop[3]),
       title="Simulated Outcomes of Biomarker and Symptom with operating characteristics",
       subtitle = )


```

```{r}
my.mean <- c(.5, .15)

my.grid <- expand.grid(Biomarker=seq(0,1,.05), Symptom=seq(0,.3,.025))
my.diag <- rbind(c(.15^2, 0), c(0, .025^2))
my.corr <- rbind(c(1, .25),c(.25, 1))
my.sigma <- sqrt(my.diag) %*% my.corr %*% sqrt(my.diag)

holdit <- bind_rows(
apply(X = matrix(1:nrow(my.grid)), MARGIN = 1, FUN = function(x) {
  my.sample <- data.frame(rmvnorm(10000, c(my.grid$Biomarker[x], my.grid$Symptom[x]), my.sigma))
names(my.sample) <- c("Biomarker", "Symptom")
my.sample <-  my.sample %>% mutate(Biomarker.result = case_when(Biomarker < .35 ~ 0,
                                                  Biomarker >=.35 & Biomarker < .50 ~ 1,
                                                  Biomarker >=.50 & Biomarker < .70 ~ 2,
                                                  TRUE >= .70 ~ 3),
                     Symptom.result = case_when(Symptom < 0.1 ~ 0,
                                                Symptom >= 0.1 & Symptom < 0.2 ~ 1,
                                                TRUE ~ 2)) %>% 
  mutate(result = case_when(
    Biomarker.result == 3 & Symptom.result != 0 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 2 ~ "Green",
    Biomarker.result == 2 & Symptom.result == 1 ~ "Yellow",
    Biomarker.result == 1 & Symptom.result != 0 ~ "Yellow", 
    .default = "Orange"
  )) %>% count(result) %>% mutate(prop = n/sum(n)) %>% select(-n) 

my.sample$result <- factor(my.sample$result, c("Green", "Yellow", "Orange") )
my.sample <- my.sample %>% mutate(Biomarker = my.grid$Biomarker[x], Symptom=my.grid$Symptom[x])
}))

holdit %>% filter(result=="Green") %>% ggplot(aes(x= Biomarker, y=Symptom, fill=prop)) + geom_tile()
holdit %>% filter(result=="Yellow") %>% ggplot(aes(x= Biomarker, y=Symptom, fill=prop)) + geom_tile()
holdit %>% filter(result=="Orange") %>% ggplot(aes(x= Biomarker, y=Symptom, fill=prop)) + geom_tile()

holdit %>% filter(Biomarker == .5) %>% ggplot(aes(x=Symptom, y= prop, color=result)) + geom_line()


```
